Measurement and Probability in Fuzzy Quantum Theories

### 2017-02-09 prop_iop.tex
prop_iop.tex Fix all fractions according "5.1. Two-line constructions"

\begin{example}[Two-coins experiment]\label{ex1} Consider an
  experiment that tosses two coins. We have four possible outcomes
  that constitute the sample space $\Omega=\{HH,HT,TH,TT\}$. There are
  16 total events including the event $\{HH,HT\}$ that the first coin
  lands heads up, the event $\{HT,TH\}$ that the two coins land on
  opposite sides, and the event $\{HT,TH,TT\}$ that at least one coin
  lands tails up. Here is a possible probability measure for these
  events:
\[
\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\pmeas(\emptyset) & = & 0\\
\pmeas(\{HH\}) & = & 1/3\\
\pmeas(\{HT\}) & = & 0\\
\pmeas(\{TH\}) & = & 2/3\\
\pmeas(\{TT\}) & = & 0\\
\pmeas(\{HH,HT\}) & = & 1/3\\
\pmeas(\{HH,TH\}) & = & 1\\
\pmeas(\{HH,TT\}) & = & 1/3
\end{array} & \begin{array}{rcl}
\pmeas(\{HT,TH\}) & = & 2/3\\
\pmeas(\{HT,TT\}) & = & 0\\
\pmeas(\{TH,TT\}) & = & 2/3\\
\pmeas(\{HH,HT,TH\}) & = & 1\\
\pmeas(\{HH,HT,TT\}) & = & 1/3\\
\pmeas(\{HH,TH,TT\}) & = & 1\\
\pmeas(\{HT,TH,TT\}) & = & 2/3\\
\pmeas(\{HH,HT,TH,TT\}) & = & 1
\end{array}\end{array}
\]
\end{example}

@@

Let's look at a concrete example. Suppose we drop a needle of length
$\ell$ onto a floor made of equally spaced parallel lines a distance
$h$ apart, where $\ell<h$. It is a known fact that the probability of
the needle crossing a line is
$\frac{2\ell}{\pi
  h}$~\cite{Buffon1777,DeMorgan1872,Hall1873,Uspensky1937}.
Consider an experimental setup consisting of a collection of $M$
identical needles of length $\ell$. We throw the $M$ needles one
needle at a time, and observe the number $M_c$ of needles that cross a
line, thus estimating the probability of a needle crossing a line to
be $\frac{M_c}{M}$. In an actual experiment with $500$ needles and the
ratio $\frac{\ell}{h}=0.75$~\cite{Hall1873}, it was found that $236$
crossed a line so the relative frequency is $0.472$ whereas the
idealized mathematical probability is $0.4774\ldots$.  In a larger
experiment with $5000$ needles and the ratio
$\frac{\ell}{h}=0.8$~\cite{Uspensky1937}, the relative frequency was
calculated to be $0.5064$ whereas the idealized mathematical
probability is $0.5092\ldots$. We see that the observed probability
approaches $\frac{2\ell}{\pi h}$ but only if \emph{larger and larger
  resources} are expended. These resource considerations suggest that
it is possible to replace the real interval $[0,1]$ with rational
numbers up to a certain precision related to the particular experiment
in question. There is clearly a connection between the number of
needles and the achievable precision: in the hypothetical experiment
with 3 needles, it is not sensible to retain 100 digits in the
expansion of $\frac{2\ell}{\pi h}$.

There is another more subtle assumption of unbounded computational
power in the experiment. We are assuming that we can always determine
with certainty whether a needle is crossing a line. But ``lines'' on
the the floor have thickness, their distance apart is not exactly $h$,
and the needles' lengths are not all absolutely equal to $\ell$.
These perturbations make the events ``fuzzy.'' Thus, in an experiment
with limited resources, it is not possible to talk about the idealized
event that exactly $M_c$ needles cross lines as this would require the
most expensive needles built to the most precise accuracy, laser
precision for drawing lines on the floor, and the most powerful
microscopes to determine if a needle does cross a line. Instead we
might talk about the event that $M_c-\delta$ needles evidently cross
lines and $M_c+\delta'$ needles plausibly cross lines where $\delta$ and
$\delta'$ are resource-dependent approximations. This fuzzy notion of
events leads to probabilities being only calculable within intervals
of confidence reflecting the certainty of events and their
plausibility. This is indeed consistent with published experiments: in
an experiment with $3204$ needles and the ratio
$\frac{\ell}{h}=0.6$~\cite{DeMorgan1872}, $1213$ needles clearly
crossed a line and $11$ needles were close enough to plausibly be
considered as crossing the line: we would express the probability in
this case as the interval
$\left[\frac{1213}{3204},\frac{1224}{3204}\right]$ expressing that we
are certain that the event has probability at least
$\frac{1213}{3204}$ but it is possible that it would have probability
$\frac{1224}{3204}$.

@@

\begin{example}[Two-coin experiment with interval probability]
\label{ex3} We split the unit interval $[0,1]$ in the following
four closed sub-intervals: $[0,0]$ which we call \imposs, $[0,\frac{1}{2}]$
which we call \unlikely, $[\frac{1}{2},1]$ which we call \likely,
and $[1,1]$ which we call \necess. Using these new values, we can
modify the probability measure of Ex.~\ref{ex1} by mapping each
numeric value to the smallest sub-interval containing it to get the
following: 
\[
\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\bar{\mu}(\emptyset) & = & \imposs\\
\bar{\mu}(\{HH\}) & = & \unlikely\\
\bar{\mu}(\{HT\}) & = & \imposs\\
\bar{\mu}(\{TH\}) & = & \likely\\
\bar{\mu}(\{TT\}) & = & \imposs\\
\bar{\mu}(\{HH,HT\}) & = & \unlikely\\
\bar{\mu}(\{HH,TH\}) & = & \necess\\
\bar{\mu}(\{HH,TT\}) & = & \unlikely
\end{array} & \begin{array}{rcl}
\bar{\mu}(\{HT,TH\}) & = & \likely\\
\bar{\mu}(\{HT,TT\}) & = & \imposs\\
\bar{\mu}(\{TH,TT\}) & = & \likely\\
\bar{\mu}(\{HH,HT,TH\}) & = & \necess\\
\bar{\mu}(\{HH,HT,TT\}) & = & \unlikely\\
\bar{\mu}(\{HH,TH,TT\}) & = & \necess\\
\bar{\mu}(\{HT,TH,TT\}) & = & \likely\\
\bar{\mu}(\{HH,HT,TH,TT\}) & = & \necess
\end{array}\end{array}
\]
Despite the absence of infinitely precise numeric information, the
probability measure is quite informative: it reveals that the second
coin is double-headed and that the first coin is biased. To understand
the $\subseteq$-condition, consider the following calculation:
\begin{eqnarray*}
\bar{\mu}(\{HH\})+\bar{\mu}(\{HT\})+\bar{\mu}(\{TH\})+\bar{\mu}(\{TT\})
&=& \imposs+\unlikely+\imposs+\likely\\
&=& \left[0,0\right]+\left[0,\frac{1}{2}\right]+\left[0,0\right]+\left[\frac{1}{2},1\right]\\
&=& \left[\frac{1}{2},\frac{3}{2}\right]
\end{eqnarray*}
If we were to equate $\bar{\mu}(\Omega)$ with the sum of the individual
probabilities, we would get that $\bar{\mu}(\Omega)=\left[\frac{1}{2},\frac{3}{2}\right]$.
However, using the fact that $\bar{\mu}(\emptyset)=\imposs$, we have
$\bar{\mu}\left(\Omega\right)=1-\bar{\mu}\left(\emptyset\right)=\necess=[1,1]$.
This interval is tighter and a better estimate for the probability
of the event $\Omega$, and of course it is contained in $[\frac{1}{2},\frac{3}{2}]$.
However it is only possible to exploit the information about the complement
when all four events are combined. Thus the $\subseteq$-condition
allows us to get an estimate for the combined event from each of its
constituents and then gather more evidence knowing the aggregate
event.
\end{example}

@@

\begin{example}[One-qubit quantum probability space] Consider a
  one-qubit Hilbert space with each event interpreted as a possible
  post-measurement state~\cite{peres1995quantum,Mermin2007,Jaeger2007}. 
  For example, the event $\proj{0}$ indicates
  that the post-measurement state will be $\ket{0}$; the probability
  of such an event depends on the current state; the event $\proj{1}$
  indicates that the post-measurement state will be $\ket{1}$; the
  event $\proj{\ps}$ where
  $\ket{\ps}=\frac{1}{\sqrt{2}}(\ket{0}+\ket{1})$ indicates that the
  post-measurement state will be $\ket{\ps}$; the event
  $\mathbb{1}=\proj{0}+\proj{1}$ indicates that the post-measurement
  state will be a linear combination of $\ket{0}$ and $\ket{1}$ and is
  clearly certain; finally the empty event $\mathbb{0}$ states that
  the post-measurement state will be the empty state and is
  impossible. As in the classical case, a probability measure is a
  function that maps events to $[0,1]$. Here is a partial
  specification of a possible probability measure that would be
  induced by a system whose current state is $\ket{0}$,
  $\mu\left(\mathbb{0}\right)=0$, $\mu\left(\mathbb{1}\right)=1$,
  $\mu\left(\proj{0}\right)=1$, $\mu\left(\proj{1}\right)=0$,
  $\mu\left(\proj{\ps}\right)=1/2$, \ldots. Note that, similarly to
  the classical case, the probability of $\mathbb{1}$ is 1 and the
  probability of collections of orthogonal events (e.g.,
  $\proj{0}+\proj{1}$) is the sum of the individual probabilities.  A
  collection of non-orthogonal events (e.g., $\proj{0}$ and
  $\proj{\ps}$) is however not even a valid event. In the classical
  example, we argued that each probability measure is uniquely
  determined by two actual coins. A similar (but much more subtle)
  argument is valid also in the quantum case. By postulates of quantum
  mechanics and Gleason's theorem, it turns out that for large enough
  quantum systems, each probability measure is uniquely determined by
  an actual quantum state as discussed next.
\end{example}

@@

\begin{example}[Three-dimensional quantum three-interval-valued
probability measure]\label{ex:three-dimensional-three-value} Given
a three dimensional Hilbert space with an orthonormal basis $\left\{ \ket{0},\ket{1},\ket{2}\right\} $.
Let $\mathscr{I}=\left\{ \necess,\imposs,\unknown\right\} $, $\ket{\ps}=\frac{1}{\sqrt{2}}(\ket{0}+\ket{1})$,
$\ket{\ps'}=\frac{1}{\sqrt{2}}(\ket{0}+\ket{2})$, and 
\begin{eqnarray*}
\bar{\mu}(\mathbb{0})=\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps})=\bar{\mu}(\proj{\ps'}) & = & \imposs\\
\bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}-\proj{0})=\bar{\mu}(\mathbb{1}-\proj{\ps})=\bar{\mu}(\mathbb{1}-\proj{\ps'}) & = & \necess\\
\bar{\mu}(P) & = & \unknown\quad\textrm{otherwise}
\end{eqnarray*}
It is straightforward to verify that $\bar{\mu}$ is a convex quantum
interval-valued probability measure. Now assume there exists a real-valued
quantum probability measure $\mu$ such that $\mu(P)\in\bar{\mu}(P)$
for every event (projection)~$P$. We derive a contradiction as follows.
Suppose $\mu(P)\in\bar{\mu}(P)$ then we must have $\mu(\proj{0})=\mu(\proj{\ps})=\mu(\proj{\ps'})=0$.
By Gleason's theorem, there is a mixed state~$\rho=\sum_{j=1}^{N}q_{j}\proj{\phi_{j}}$
such that $\mu\left(P\right)=\sum_{j=1}^{N}q_{j}\ip{\phi_{j}}{P\phi_{j}}$,
where $\sum_{j=1}^{N}q_{j}=1$ and $q_{j}>0$. However, no pure state~$\ket{\phi}$
can satisfy $\ip{\phi}{0}=\ip{\phi}{\ps}=\ip{\phi}{\ps'}=0$; a contradiction.
\end{example}

@@

Looking closely at the example above, we might think that the probability
measure $\bar{\mu}$ is induced by the state $\ket{2}$ because $\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps})=\imposs$,
or we might think it is induced by the state $\ket{1}$ because $\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps'})=\imposs$,
or yet we might think it is induced by the mixed state $\frac{\mathbb{1}}{3}$
because $\bar{\mu}(\proj{\phi})\ne\necess$ for all $\ket{\phi}$.
Each of these possibilities is compatible with some but not all of
the observations. All is not lost however: if the observations are
made more precise, we conjecture that some of the inconsistencies
disappear. A partial proof of this conjecture is the following example
which refines the previous convex probability measure by adding more
precise intervals.

\begin{example}[Three-dimensional quantum four-interval-valued 
  probability measure]\label{ex:three-dimensional-4-value}
  Given a three dimensional Hilbert space with an orthonormal basis
  $\left\{ \ket{0},\ket{1},\ket{2}\right\} $.  Let \emph{
  }$\mathscr{I}=\left\{ \imposs,\unlikely,\likely,\necess\right\} $,
  $\ket{\ps}=\frac{1}{\sqrt{2}}(\ket{0}+\ket{1})$, and
  $\ket{\minus}=\frac{1}{\sqrt{2}}(\ket{0}-\ket{1})$. The definition of
  $\bar{\mu}$ below refers to figure~\ref{fig:three-dimensional-4-value}
  which plots the 1-dimensional projectors:
\begin{enumerate}
\item Let:
\begin{eqnarray*}
\bar{\mu}(\mathbb{0})=\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps}) 
  &=& \imposs \\
\bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}-\proj{0})=\bar{\mu}(\mathbb{1}-\proj{\ps}) 
  &=& \necess
\end{eqnarray*}
where $\ket{0}$ and $\ket{\ps}$ are plotted as the red and green dotted
vectors, respectively.
\item The red and green circles are are the states orthogonal to
  $\ket{0}$ and $\ket{\ps}$, respectively, and can be parametrized as
  $\ket{0_{\theta,\gamma}^{\perp}}=\rme^{\rmi\gamma}\sin\theta\ket{1}+\cos\theta\ket{2}$
  and
  $\ket{\ps_{\theta,\gamma}^{\perp}}=-\rme^{\rmi\gamma}\sin\theta\ket{\minus}+\cos\theta\ket{2}$,
  where $0\le\theta\le\frac{\pi}{2}$ and $0\le\gamma<2\pi$. The dotted
  half of those states need special treatment, i.e., whenever
  $0\le\theta<\frac{\pi}{2}$ and $0\le\gamma<\pi$, we define
\begin{eqnarray*}
\bar{\mu}\left(\proj{0_{\theta,\gamma}^{\perp}}\right)=\bar{\mu}\left(\proj{\ps_{\theta,\gamma}^{\perp}}\right)
  &=& \likely \\
\bar{\mu}\left(\mathbb{1}-\proj{0_{\theta,\gamma}^{\perp}}\right)=
  \bar{\mu}\left(\mathbb{1}-\proj{\ps_{\theta,\gamma}^{\perp}}\right) 
  &=& \unlikely
\end{eqnarray*}
\item Otherwise, $\bar{\mu}(\proj{\psi})=\unlikely$ and $\bar{\mu}(\mathbb{1}-\proj{\psi})=\likely$. 
\end{enumerate}
It is straightforward but tedious to check that $\bar{\mu}$ is a
quantum interval-valued probability measure. Although we have not
verified that $\bar{\mu}$ is convex, the following argument establishes
that is has an empty core. Assume there is a real-valued probability
measure satisfying $\mu_{\rho}^{B}(P)\in\bar{\mu}(P)$ for all $P\in\events$.
Because $\mu_{\rho}^{B}(\proj{0})\in\bar{\mu}(\proj{0})=\imposs$
and $\mu_{\rho}^{B}(\proj{\ps})\in\bar{\mu}(\proj{\ps})=\imposs$,
we must have $\mu_{\rho}^{B}(\proj{0})=\mu_{\rho}^{B}(\proj{\ps})=0$
so that $\mu_{\rho}^{B}=\mu_{\ket{2}}^{B}$. However, 
\[
\mu_{\ket{2}}^{B}(\proj{2})=1\notin\unlikely=\bar{\mu}(\proj{2})\textrm{ .}
\]
This measure however is ``better'' than the previous one in the
sense that it might only be induced by $\ket{2}$ or the density matrix
$\frac{\mathbb{1}}{3}$, but not by $\ket{1}$. \end{example}

@@

\begin{definition}Given a vector~$\ket{\psi}\in S^{2}$ with a specified
north pole~$\ket{0}\in S^{2}$, then $\angle\left(\ket{\psi}\right)=\arccos\ip{0}{\psi}$
denotes the latitude of $\ket{\psi}$. Conversely, $\angle^{-1}\left(I\right)=\set{\ket{\psi}\in S^{2}}{\angle\left(\ket{\psi}\right)\in I}$.
For example, the north pole has the latitude $0$, the equator has
the the latitude $\frac{\pi}{2}$, and $\angle^{-1}\left(\left[0,\theta_{0}\right]\right)=\set{\ket{\psi}\in S^{2}}{\theta_{0}\le\angle\left(\ket{\psi}\right)\le\frac{\pi}{2}}$
contains all states whose latitude is larger than $\theta_{0}$.\end{definition}

>>>>

\begin{example}[Two-coins experiment]\label{ex1} Consider an
  experiment that tosses two coins. We have four possible outcomes
  that constitute the sample space $\Omega=\{HH,HT,TH,TT\}$. There are
  16 total events including the event $\{HH,HT\}$ that the first coin
  lands heads up, the event $\{HT,TH\}$ that the two coins land on
  opposite sides, and the event $\{HT,TH,TT\}$ that at least one coin
  lands tails up. Here is a possible probability measure for these
  events:
\[
\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\pmeas(\emptyset) & = & 0\\
\pmeas(\{HH\}) & = & \frac{1}{3}\\
\pmeas(\{HT\}) & = & 0\\
\pmeas(\{TH\}) & = & \frac{2}{3}\\
\pmeas(\{TT\}) & = & 0\\
\pmeas(\{HH,HT\}) & = & \frac{1}{3}\\
\pmeas(\{HH,TH\}) & = & 1\\
\pmeas(\{HH,TT\}) & = & \frac{1}{3}
\end{array} & \begin{array}{rcl}
\pmeas(\{HT,TH\}) & = & \frac{2}{3}\\
\pmeas(\{HT,TT\}) & = & 0\\
\pmeas(\{TH,TT\}) & = & \frac{2}{3}\\
\pmeas(\{HH,HT,TH\}) & = & 1\\
\pmeas(\{HH,HT,TT\}) & = & \frac{1}{3}\\
\pmeas(\{HH,TH,TT\}) & = & 1\\
\pmeas(\{HT,TH,TT\}) & = & \frac{2}{3}\\
\pmeas(\{HH,HT,TH,TT\}) & = & 1
\end{array}\end{array}
\]
\end{example}

@@

Let's look at a concrete example. Suppose we drop a needle of length
$\ell$ onto a floor made of equally spaced parallel lines a distance
$h$ apart, where $\ell<h$. It is a known fact that the probability of
the needle crossing a line is
$2\ell/\left(\pi h\right)$~\cite{Buffon1777,DeMorgan1872,Hall1873,Uspensky1937}.
Consider an experimental setup consisting of a collection of $M$
identical needles of length $\ell$. We throw the $M$ needles one
needle at a time, and observe the number $M_c$ of needles that cross a
line, thus estimating the probability of a needle crossing a line to
be $M_c/M$. In an actual experiment with $500$ needles and the
ratio $\ell/h=0.75$~\cite{Hall1873}, it was found that $236$
crossed a line so the relative frequency is $0.472$ whereas the
idealized mathematical probability is $0.4774\ldots$.  In a larger
experiment with $5000$ needles and the ratio
$\ell/h=0.8$~\cite{Uspensky1937}, the relative frequency was
calculated to be $0.5064$ whereas the idealized mathematical
probability is $0.5092\ldots$. We see that the observed probability
approaches $2\ell/\left(\pi h\right)$ but only if \emph{larger and larger
  resources} are expended. These resource considerations suggest that
it is possible to replace the real interval $[0,1]$ with rational
numbers up to a certain precision related to the particular experiment
in question. There is clearly a connection between the number of
needles and the achievable precision: in the hypothetical experiment
with 3 needles, it is not sensible to retain 100 digits in the
expansion of $2\ell/\left(\pi h\right)$.

There is another more subtle assumption of unbounded computational
power in the experiment. We are assuming that we can always determine
with certainty whether a needle is crossing a line. But ``lines'' on
the the floor have thickness, their distance apart is not exactly $h$,
and the needles' lengths are not all absolutely equal to $\ell$.
These perturbations make the events ``fuzzy.'' Thus, in an experiment
with limited resources, it is not possible to talk about the idealized
event that exactly $M_c$ needles cross lines as this would require the
most expensive needles built to the most precise accuracy, laser
precision for drawing lines on the floor, and the most powerful
microscopes to determine if a needle does cross a line. Instead we
might talk about the event that $M_c-\delta$ needles evidently cross
lines and $M_c+\delta'$ needles plausibly cross lines where $\delta$ and
$\delta'$ are resource-dependent approximations. This fuzzy notion of
events leads to probabilities being only calculable within intervals
of confidence reflecting the certainty of events and their
plausibility. This is indeed consistent with published experiments: in
an experiment with $3204$ needles and the ratio
$\ell/h=0.6$~\cite{DeMorgan1872}, $1213$ needles clearly
crossed a line and $11$ needles were close enough to plausibly be
considered as crossing the line: we would express the probability in
this case as the interval
$\left[\frac{1213}{3204},\frac{1224}{3204}\right]$ expressing that we
are certain that the event has probability at least
$\frac{1213}{3204}$ but it is possible that it would have probability
$\frac{1224}{3204}$.

@@

\begin{example}[Two-coin experiment with interval probability]
\label{ex3} We split the unit interval $[0,1]$ in the following
four closed sub-intervals: $[0,0]$ which we call \imposs, $[0,\frac{1}{2}]$
which we call \unlikely, $[\frac{1}{2},1]$ which we call \likely,
and $[1,1]$ which we call \necess. Using these new values, we can
modify the probability measure of example~\ref{ex1} by mapping each
numeric value to the smallest sub-interval containing it to get the
following: 
\[
\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\bar{\mu}(\emptyset) & = & \imposs\\
\bar{\mu}(\{HH\}) & = & \unlikely\\
\bar{\mu}(\{HT\}) & = & \imposs\\
\bar{\mu}(\{TH\}) & = & \likely\\
\bar{\mu}(\{TT\}) & = & \imposs\\
\bar{\mu}(\{HH,HT\}) & = & \unlikely\\
\bar{\mu}(\{HH,TH\}) & = & \necess\\
\bar{\mu}(\{HH,TT\}) & = & \unlikely
\end{array} & \begin{array}{rcl}
\bar{\mu}(\{HT,TH\}) & = & \likely\\
\bar{\mu}(\{HT,TT\}) & = & \imposs\\
\bar{\mu}(\{TH,TT\}) & = & \likely\\
\bar{\mu}(\{HH,HT,TH\}) & = & \necess\\
\bar{\mu}(\{HH,HT,TT\}) & = & \unlikely\\
\bar{\mu}(\{HH,TH,TT\}) & = & \necess\\
\bar{\mu}(\{HT,TH,TT\}) & = & \likely\\
\bar{\mu}(\{HH,HT,TH,TT\}) & = & \necess
\end{array}\end{array}
\]
Despite the absence of infinitely precise numeric information, the
probability measure is quite informative: it reveals that the second
coin is double-headed and that the first coin is biased. To understand
the $\subseteq$-condition, consider the following calculation:
\begin{eqnarray*}
& & \bar{\mu}(\{HH\})+\bar{\mu}(\{HT\})+\bar{\mu}(\{TH\})+\bar{\mu}(\{TT\})\\
&=& \imposs+\unlikely+\imposs+\likely\\
&=& \left[0,0\right]+\left[0,\frac{1}{2}\right]+\left[0,0\right]+\left[\frac{1}{2},1\right]\\
&=& \left[\frac{1}{2},\frac{3}{2}\right]
\end{eqnarray*}
If we were to equate $\bar{\mu}(\Omega)$ with the sum of the individual
probabilities, we would get that $\bar{\mu}(\Omega)=\left[\frac{1}{2},\frac{3}{2}\right]$.
However, using the fact that $\bar{\mu}(\emptyset)=\imposs$, we have
$\bar{\mu}\left(\Omega\right)=1-\bar{\mu}\left(\emptyset\right)=\necess=[1,1]$.
This interval is tighter and a better estimate for the probability
of the event $\Omega$, and of course it is contained in $[\frac{1}{2},\frac{3}{2}]$.
However it is only possible to exploit the information about the complement
when all four events are combined. Thus the $\subseteq$-condition
allows us to get an estimate for the combined event from each of its
constituents and then gather more evidence knowing the aggregate
event.
\end{example}

@@

\begin{example}[One-qubit quantum probability space] Consider a
  one-qubit Hilbert space with each event interpreted as a possible
  post-measurement state~\cite{peres1995quantum,Mermin2007,Jaeger2007}. 
  For example, the event $\proj{0}$ indicates
  that the post-measurement state will be $\ket{0}$; the probability
  of such an event depends on the current state; the event $\proj{1}$
  indicates that the post-measurement state will be $\ket{1}$; the
  event $\proj{\ps}$ where
  $\ket{\ps}=\left(\ket{0}+\ket{1}\right) / \sqrt{2}$ indicates that the
  post-measurement state will be $\ket{\ps}$; the event
  $\mathbb{1}=\proj{0}+\proj{1}$ indicates that the post-measurement
  state will be a linear combination of $\ket{0}$ and $\ket{1}$ and is
  clearly certain; finally the empty event $\mathbb{0}$ states that
  the post-measurement state will be the empty state and is
  impossible. As in the classical case, a probability measure is a
  function that maps events to $[0,1]$. Here is a partial
  specification of a possible probability measure that would be
  induced by a system whose current state is $\ket{0}$,
  $\mu\left(\mathbb{0}\right)=0$, $\mu\left(\mathbb{1}\right)=1$,
  $\mu\left(\proj{0}\right)=1$, $\mu\left(\proj{1}\right)=0$,
  $\mu\left(\proj{\ps}\right)=\frac{1}{2}$, \ldots. Note that, similarly to
  the classical case, the probability of $\mathbb{1}$ is 1 and the
  probability of collections of orthogonal events (e.g.,
  $\proj{0}+\proj{1}$) is the sum of the individual probabilities.  A
  collection of non-orthogonal events (e.g., $\proj{0}$ and
  $\proj{\ps}$) is however not even a valid event. In the classical
  example, we argued that each probability measure is uniquely
  determined by two actual coins. A similar (but much more subtle)
  argument is valid also in the quantum case. By postulates of quantum
  mechanics and Gleason's theorem, it turns out that for large enough
  quantum systems, each probability measure is uniquely determined by
  an actual quantum state as discussed next.
\end{example}

@@

\begin{example}[Three-dimensional quantum three-interval-valued
probability measure]\label{ex:three-dimensional-three-value} Given
a three dimensional Hilbert space with an orthonormal basis $\left\{ \ket{0},\ket{1},\ket{2}\right\} $.
Let $\mathscr{I}=\left\{ \necess,\imposs,\unknown\right\} $, $\ket{\ps}=\left(\ket{0}+\ket{1}\right) / \sqrt{2}$,
$\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$, and 
\begin{eqnarray*}
\bar{\mu}(\mathbb{0}) & = & \bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps})=\bar{\mu}(\proj{\ps'})=\imposs\\
\bar{\mu}(\mathbb{1}) & = & \bar{\mu}(\mathbb{1}-\proj{0})=\bar{\mu}(\mathbb{1}-\proj{\ps})=\bar{\mu}(\mathbb{1}-\proj{\ps'})=\necess\\
\bar{\mu}(P) & = & \unknown\quad\textrm{otherwise}
\end{eqnarray*}
It is straightforward to verify that $\bar{\mu}$ is a convex quantum
interval-valued probability measure. Now assume there exists a real-valued
quantum probability measure $\mu$ such that $\mu(P)\in\bar{\mu}(P)$
for every event (projection)~$P$. We derive a contradiction as follows.
Suppose $\mu(P)\in\bar{\mu}(P)$ then we must have $\mu(\proj{0})=\mu(\proj{\ps})=\mu(\proj{\ps'})=0$.
By Gleason's theorem, there is a mixed state~$\rho=\sum_{j=1}^{N}q_{j}\proj{\phi_{j}}$
such that $\mu\left(P\right)=\sum_{j=1}^{N}q_{j}\ip{\phi_{j}}{P\phi_{j}}$,
where $\sum_{j=1}^{N}q_{j}=1$ and $q_{j}>0$. However, no pure state~$\ket{\phi}$
can satisfy $\ip{\phi}{0}=\ip{\phi}{\ps}=\ip{\phi}{\ps'}=0$; a contradiction.
\end{example}

@@

Looking closely at the example above, we might think that the probability
measure $\bar{\mu}$ is induced by the state $\ket{2}$ because $\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps})=\imposs$,
or we might think it is induced by the state $\ket{1}$ because $\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps'})=\imposs$,
or yet we might think it is induced by the mixed state $\mathbb{1} / 3$
because $\bar{\mu}(\proj{\phi})\ne\necess$ for all $\ket{\phi}$.
Each of these possibilities is compatible with some but not all of
the observations. All is not lost however: if the observations are
made more precise, we conjecture that some of the inconsistencies
disappear. A partial proof of this conjecture is the following example
which refines the previous convex probability measure by adding more
precise intervals.

\begin{example}[Three-dimensional quantum four-interval-valued 
  probability measure]\label{ex:three-dimensional-4-value}
  Given a three dimensional Hilbert space with an orthonormal basis
  $\left\{ \ket{0},\ket{1},\ket{2}\right\} $.  Let \emph{
  }$\mathscr{I}=\left\{ \imposs,\unlikely,\likely,\necess\right\} $,
  $\ket{\ps}=\left(\ket{0}+\ket{1}\right) / \sqrt{2}$, and
  $\ket{\minus}=\left(\ket{0}-\ket{1}\right) / \sqrt{2}$. The definition of
  $\bar{\mu}$ below refers to figure~\ref{fig:three-dimensional-4-value}
  which plots the 1-dimensional projectors:
\begin{enumerate}
\item Let:
\begin{eqnarray*}
\bar{\mu}(\mathbb{0}) &=& \bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps})=\imposs \\
\bar{\mu}(\mathbb{1}) &=& \bar{\mu}(\mathbb{1}-\proj{0})=\bar{\mu}(\mathbb{1}-\proj{\ps})=\necess
\end{eqnarray*}
where $\ket{0}$ and $\ket{\ps}$ are plotted as the red and green dotted
vectors, respectively.
\item The red and green circles are are the states orthogonal to
  $\ket{0}$ and $\ket{\ps}$, respectively, and can be parametrized as
  $\ket{0_{\theta,\gamma}^{\perp}}=\rme^{\rmi\gamma}\sin\theta\ket{1}+\cos\theta\ket{2}$
  and
  $\ket{\ps_{\theta,\gamma}^{\perp}}=-\rme^{\rmi\gamma}\sin\theta\ket{\minus}+\cos\theta\ket{2}$,
  where $0\le\theta\le\pi / 2$ and $0\le\gamma<2\pi$. The dotted
  half of those states need special treatment, i.e., whenever
  $0\le\theta<\pi / 2$ and $0\le\gamma<\pi$, we define
\begin{eqnarray*}
\bar{\mu}\left(\proj{0_{\theta,\gamma}^{\perp}}\right) &=& \bar{\mu}\left(\proj{\ps_{\theta,\gamma}^{\perp}}\right)=\likely \\
\bar{\mu}\left(\mathbb{1}-\proj{0_{\theta,\gamma}^{\perp}}\right) &=& 
  \bar{\mu}\left(\mathbb{1}-\proj{\ps_{\theta,\gamma}^{\perp}}\right)=\unlikely
\end{eqnarray*}
\item Otherwise, $\bar{\mu}(\proj{\psi})=\unlikely$ and $\bar{\mu}(\mathbb{1}-\proj{\psi})=\likely$. 
\end{enumerate}
It is straightforward but tedious to check that $\bar{\mu}$ is a
quantum interval-valued probability measure. Although we have not
verified that $\bar{\mu}$ is convex, the following argument establishes
that is has an empty core. Assume there is a real-valued probability
measure satisfying $\mu_{\rho}^{B}(P)\in\bar{\mu}(P)$ for all $P\in\events$.
Because $\mu_{\rho}^{B}(\proj{0})\in\bar{\mu}(\proj{0})=\imposs$
and $\mu_{\rho}^{B}(\proj{\ps})\in\bar{\mu}(\proj{\ps})=\imposs$,
we must have $\mu_{\rho}^{B}(\proj{0})=\mu_{\rho}^{B}(\proj{\ps})=0$
so that $\mu_{\rho}^{B}=\mu_{\ket{2}}^{B}$. However, 
\[
\mu_{\ket{2}}^{B}(\proj{2})=1\notin\unlikely=\bar{\mu}(\proj{2})\textrm{ .}
\]
This measure however is ``better'' than the previous one in the
sense that it might only be induced by $\ket{2}$ or the density matrix
$\mathbb{1} / 3$, but not by $\ket{1}$. \end{example}

@@

\begin{definition}Given a vector~$\ket{\psi}\in S^{2}$ with a specified
north pole~$\ket{0}\in S^{2}$, then $\angle\left(\ket{\psi}\right)=\arccos\ip{0}{\psi}$
denotes the latitude of $\ket{\psi}$. Conversely, $\angle^{-1}\left(I\right)=\set{\ket{\psi}\in S^{2}}{\angle\left(\ket{\psi}\right)\in I}$.
For example, the north pole has the latitude $0$, the equator has
the the latitude $\pi / 2$, and $\angle^{-1}\left(\left[0,\theta_{0}\right]\right)=\set{\ket{\psi}\in S^{2}}{\theta_{0}\le\angle\left(\ket{\psi}\right)\le\pi / 2}$
contains all states whose latitude is larger than $\theta_{0}$.\end{definition}
